# Data Exploration

Exploratory data analysis (EDA) is the first step in the getting insights from a dataset. In this section we will see look at our cleaned dataset and perform some summary statistics and visualizations to see what would be a good destinations for our travellers to head to. 

## Summary statistics

We can make use of the **summary( )** function that was introduced in Chapter 4 to determine the mean and mean costs for **meals** and **one way ticket** costs. We can make use of indexing to select just these two columns. 

```{r summary}
summary(travel_full_clean_subset[3:4])
```
We can see that the **median** costs for a meal \$10.53 CAD, and the **mean** costs is \$11.84 CAD. The lowest priced meals are shown as **min** \$2.05 CAD, and the most expensive is **max** \$26.72 CAD. 

We can use this information to filter out countries that we want to visit based on food or ticket costs. Lets consider filtering out food by the interquartile range (**IQR**). Knowing that the 1st quartile is \$6.32 CAD and the 3rd quartile is \$17.49 we can use the **filter( )** function and the **between( )** function to select only those countries in this range.

```{r filter meal low}
travel_full_clean_subset %>%
  filter(between(Meal_Inexpensive_Restaurant,6.32,17.49)) %>%
  select(c(Country,Meal_Inexpensive_Restaurant)) %>%
  arrange(Meal_Inexpensive_Restaurant) %>%
  head(5)
```

If we want to rank the previous countries based on meals from least expensive to most expensive we can use the **arrange( )** function. If we would like to go from most expensive to least expensive, we can add the **desc( )** function. 

<h3> Note </h3>

> We can also use the **tail( )** function instead of the **head( )** function to get the most expensive.


```{r filter meal high}
travel_full_clean_subset %>%
  filter(between(Meal_Inexpensive_Restaurant,6.32,17.49)) %>%
  select(c(Country,Meal_Inexpensive_Restaurant)) %>%
  arrange(desc(Meal_Inexpensive_Restaurant)) %>%
  head(5)
```

On the lower meal price end we have 

* Ethiopia
 
* Jamaica

* Ukraine

* Jordon

* Albania


and the higher end we have 

* Puerto Rico

* Sweden

* Latvia

* Solenia

* Georgia

We combining what we the data wrangling from the previous section, we can determine what the average number of arrivals for each country is from the lat 5 years using the functions **mutate( )** and **arrange( )**
```{r most travelled 5year}
travel_full_clean_subset %>%
  mutate(
    mean5year = rowMeans(select(., "2015":"2020"))) %>%
  select(c(Country,mean5year)) %>%
  arrange(desc(mean5year)) %>%
  head(5)

```

## Standard deviation and variance 

We can easily calculate the dispersion of the data using standard deviaton and variance that are built in functions with R. 

### Standard deviation 

Standard deviation explains how far away a group of numbers is from the mean. If we wanted to find the standard deviation for our meals, we can just the **sd( )** function.

<h3> Note </h3>

> The **sd( )** function will generate the sample standard deviation. 

> To calculate the population standard deivation we would need to multiple the standard deviation by sqrt((n-1)/n) * sd(x)

```{r standard deivation meal}
sd(travel_full_clean_subset$Meal_Inexpensive_Restaurant)
```

### Varience 

Variance is how spread out the data is around the mean. To calculate variance the meals variable we can use the **var( )** function. 

<h3> Note </h3>

> The **var( )** function will generate the sample varience. 
> If you ever need to calculate the population varience we would need to multiple the varience by (n-1)/n * var(x)

```{r varience meal}
var(travel_full_clean_subset$Meal_Inexpensive_Restaurant)
```

Although it is easy to perform these functions, it is important to understand what you are calculating and if it is appropriate to use that calculation. 

